---
title: "Statistical assignment 1"
author: "Alicia Rey-Herme (660004693)"
date: "24 January 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please do not forget to add your name and student number above!

## Open data (20 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}

# Install tidyverse
library(tidyverse)

# Loading in data. Data folder was renamed to SN6614 for efficiency
Data <- read_tsv("C:/Users/alici/Documents/DA3/data/SN6614/tab/ukhls_w8/h_indresp.tab")

```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

```{r}

# Choosing which variables to keep (See Above for identification)
Data <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)

```

## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71).We will not do this in this assignment.

However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain.

```{r}

# Filtering h_memorig to keep only the original Understanding Society Sample, "ukhls gb 2009-10" (Coded as 1)
Data <- Data %>%
        filter(h_memorig == 1)

```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/wave/8/datafile/h_indresp .

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.
2) Recode sex into a character vector with the values "male" or "female".
3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

```{r}

class(Data$h_sex_dv)

# Recoding EU Variable to a Binary
Data <- Data %>%
        mutate(EU = case_when(
              h_eumem < 1 ~ NA_real_ ,  # All missing values -> NA
              h_eumem == 1 ~ 1,  # All Remain values (1) -> 1
              h_eumem == 2 ~ 0  # All Leave values (2) -> 0
              )
          ) %>%
# Recoding Sex
      mutate(Sex = recode(h_sex_dv,
                          `1` = "male",  # "1" numeric value -> character "male" value
                          `2` = "female",  # "2" numeric value -> character "female" value
                          .default = NA_character_  # Any other value -> NA
                    )
             ) %>% 
# Recoding Age
        mutate(AgeRange = case_when(
                between(h_age_dv, 16, 25) ~ "16 - 25",  
                between(h_age_dv, 26, 40) ~ "26 - 40",  
                between(h_age_dv, 41, 55) ~ "41 - 55",  
                between(h_age_dv, 56, 70) ~ "56 - 70",  
                h_age_dv > 70 ~ "Over 70"  
          )
        )
                
# check the use of between by typing "? between" in the console
?between
        
```

## Summarise data (10 points)

Calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}

# Calculating respondent's support of Remain and Leave
Data %>%
        filter(!is.na(EU)) %>%  # Eliminate NA values in EU
        count(EU) %>%  # Count number or 0 and 1 values in EU
        mutate(Percentage = n / sum(n) )  # Calcuate percentage
                                    
```

At 54.35%, respondents in the sample appear to have supported the Remain movement more than Leave. This is contrary to the result of the 2016 referendum, where Leave won with 51.89% of the vote rather than the 45.65% suggested by the sample results. The margin between the two fields of support is further a lot narrower in the 2016 results than in the sample size, at ~4% rather than ~10%.

This could have been a result of a skewed sample size. Although stratified and clustered, the response rate of ~60% could suggest that a smaller proportion of Leave voters replied to the survey than that of Remain. Furthermore, the comparison could suggest a change in public opinion over time, a lower voter turn out of Remain supporters than that of Leave, or that people could have been dishonest with their political views when filling out the survey, so that the support recorded did not match the final result, 



## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.

By sex:

```{r}

# Group by Sex, remove NAs, and calculate percentage through mean()
Data %>%
        group_by(Sex) %>%  # Categorise by Sex
        drop_na() %>%  # Eliminate all NAs
        summarise(meanEU = mean(EU, na.rm = TRUE))  # Calculate Probability through mean()
  
```

By age groups:

```{r}

# Group by Age, remove NAs, and calculate percentage through mean()
Data %>%
        group_by(AgeRange) %>% # Categorise by AgeRange
        drop_na() %>%
        summarise(meanEU = mean(EU, na.rm = TRUE)) 

```

By sex and age group:

```{r}

# Group by Age and Sex, remove NAs, and calculate percentage through mean()
Data %>%
        group_by(AgeRange, Sex) %>%  # Categorise by Age Range and Sex
        drop_na() %>%
        summarise(meanEU = mean(EU, na.rm = TRUE))
```

It appears that females were more likely to support Remain than males, a trend that is consistent in all age brackets. The younger the respondent the more likely it was that they supported Remain. In fact, it is only in the 56-70 age bracket and above that a larger proportion of respondents supported Leave than Remain. This is with exclusion of females in the 56-70 age bracket, who still were marginally more likely to support Remain.