---
title: "Test statistical assignment"
author: "Alicia Rey-Herme (660004693)"
date: "22 January 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Reading data (40 points)

1) The Understanding Society Waves 1-8 User Guide: https://www.understandingsociety.ac.uk/sites/default/files/downloads/documentation/mainstage/user-guides/mainstage-user-guide.pdf
2) The youth self-completion questionnaire from Wave 8: https://www.understandingsociety.ac.uk/sites/default/files/downloads/documentation/mainstage/questionnaire/wave-8/W8-youth-questionnaire.pdf
3) The codebook for the file: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/wave/8/datafile/h_youth

```{r}

# Install tidyverse
library(tidyverse)

# Loading in data. Data folder was renamed to SN6614 for efficiency
Data <- read_tsv("C:/Users/alici/Documents/DA3/data/SN6614/tab/ukhls_w8/h_youth.tab")

```


## Tabulate variables (10 points)

In the survey children were asked the following question: "Do you have a social media profile or account on any sites or apps?". In this assignment we want to explore how the probability of having an account on social media depends on children's age and gender.

```{r}
# Column Names:
# h_sex_dv <- Respondent's Gender (Derived Variables) | Male (1), Female (2), NA (-9)
# h_age_dv <- Respondent's Age (Derived Variables) | Ages 10-16
# h_ypsocweb <- Respondent's answer to "Do you have a social media profile or account on any sites or apps? | Yes (1), No (2), NA (-9)


#Individual Freq. Tables
table(Data$h_sex_dv)
table(Data$h_age_dv)
table(Data$h_ypsocweb)

```

## Recode variables (10 points)

```{r}
# Create a new binary variable (SocialMedia) by recoding variable h_ypsocweb so that 1 means "yes", 0 means "no", missing values as NA
Data$SocialMedia[Data$h_ypsocweb=="1"] <- 1
Data$SocialMedia[Data$h_ypsocweb=="2"] <- 0
Data$SocialMedia[Data$h_ypsocweb=="-9"] <- NA

# Ensure variable SocialMedia is numeric 
class(Data$SocialMedia) <- "numeric"

# Create a new variable (Gender) by recoding variable h_sex_dv with the values "male" and "female"
Data$Gender[Data$h_sex_dv==1] <- "male"
Data$Gender[Data$h_sex_dv==2] <- "female"

# Rename h_age_dv to Age for consistency
Data$Age <- Data$h_age_dv

```

## Calculate means (10 points)

```{r}

# Calculating probability for each pairing of Age and Gender
   Data %>%
   group_by(Age, Gender) %>%  # Categorise by Age and Gender
   filter(Age != 16) %>%  # Filter out respondents aged 16
   summarise(Probability=mean(SocialMedia, na.rm = TRUE))  # Using mean() to calculate probability

```

## Write short interpretation (10 points)

For both male and female respondents, the probability of having a social media account increases with age. At every age value, females have a higher probability of having an account of social media than males. Aside from 10 year old males, who have a probability of 0.48%, all categories of respondents had over a 50% likelihood of owning a social media account, suggesting that the majority of respondents own an account.

## Visualise results (20 points)

```{r}

# Plot a bar graph using aforementioned function
Data %>%
   group_by(Age, Gender) %>%
   summarise(Probability=mean(SocialMedia, na.rm = TRUE)) %>%
   filter(Age != 16) %>%
    ggplot(aes(fill=Gender, y=Probability, x=Age)) +  # Plotting the graph
    geom_bar(position="dodge", stat="identity") +  # Choosing bar graph and its' properties
   ggtitle("Probability of owning a Social Media Account given Gender and Age")  # Add title
```

A bar plot would be most appropriate for visualising the data as it can compare probability values between female and male respondents as well as track the change over age.

## Conclusion

This is a test formative assignment and the mark will not count towards your final mark. If you cannot answer any of the questions above this is fine -- we are just starting this module! However, please do submit this assignment in any case to make sure you understand the procedure, it works correctly and we do not have any problems with summative assignments later.
